---
title: "Basic plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(icorp26)
library(easystats)
```

We can also make some simple bar charts to 
display these. (There are many types of plots but these 
are to get started.)


```{r}
data_tabulate(rss1$HIS_GENERAL) |> plot() 
```

That's a mess! 

Fortunately, we can use some options from the 
`ggplot2` package and some extra options.

```{r}
library(ggplot2)
```

```{r}
data_tabulate(rss1$HIS_GENERAL) |> 
     plot(label_values = FALSE, error_bar = FALSE,
          show_na = "never") + 
     coord_flip() +
     labs(title = "Overall health")

data_tabulate(rss1$PAY_PAYWORRY) |>
     plot(label_values = FALSE, error_bar = FALSE,
          show_na = "never") + 
     coord_flip() +
     labs(title = "Worried about pay if sick/injured") 

data_tabulate(rss1$MARSTAT) |> 
     plot(label_values = FALSE, error_bar = FALSE,
          show_na = "never") + 
     coord_flip() +
     labs(title = "Marital Status")
```

A histogram is one way to present a 
distribution of a numeric variable is using
a histogram.  For this example I will
use full ggplot code. 

```{r}
ggplot(rss1, aes(as.numeric(P_INCOME_R))) +
     geom_histogram()
```

Here is an example of the bar plot
using ggplot.

```{r}
ggplot(rss1, aes(MARSTAT)) +
     geom_bar() +
     coord_flip() +
     labs(title = "Marital Status")
```



They definitely are more complex, but you also have
a huge amount of flexibility. You can change
colors,shapes, sizes and more.







