---
title: "ICORP Cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Significance tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(icorp26)
library(easystats)
library(ggplot2)
library(gtsummary)
library(table1)
```
This document shows how to do a number of tasks
you may want to do in your poster.

Remember  that have _data frames_ and then _variables_ within the data frame.
We generally refer to one variable in a dataframe as
`data_frame_name$variable_name`. Both must be spelled exactly
correctly and with the right upper and lower case letters.

You also need to make sure that you are using numeric variables
when you need numbers. 

Get means for groups with an F test.
```{r}
means_by_group(to_numeric(rss1$SUN_USEFACE), 
               by =rss1$P_GENDER)
```

Plot the means by group with an F test

```{r}
means_by_group(to_numeric(rss1$SUN_USEFACE), 
               by =rss1$P_GENDER) |> plot()
```

Make a frequency table

```{r}
data_tabulate(rss1$SUN_CLOUDY, metrics = c("N", "valid"), 
              remove_na = TRUE)
```

Make a cross tabulation

```{r}
data_tabulate(rss1$SUN_USEFACE,rss1$P_GENDER , 
              remove_na = TRUE, proportions = "column") |>
     print_md()
```

Visualize a cross tab

```{r}
data_tabulate(rss1$SUN_USEFACE,rss1$P_GENDER , 
              remove_na = TRUE, proportions = "column") |> plot() +
     labs(title = "Use of Sun Screen on face by Gender",
          x = "Gender", y = "How often")
```

Get the chi square for a cross tab.

```{r}
data_tabulate(rss1$SUN_USEFACE,rss1$P_GENDER , 
              remove_na = TRUE, proportions = "column") |> 
     data_chisq()
```

Basic bar chart of one variable.

```{r}
distribution_bar(rss1$SUN_CLOUDY)
```


Compare a lot of categorical/nominal variables
that all have the same responses 
in a compact way. There  are alot ofpossivllwe ways to
do this.

```{r}
library(dplyr)
library(tidyr)
rss1 |> select("SUN_NOHARM", "SUN_CLOUDY") |>
     pivot_longer(cols = c("SUN_NOHARM", "SUN_CLOUDY"), 
                  names_to = "variable",  
                  values_to = "response") |>
     data_tabulate( "variable", by = "response", 
                    proportions = "row", remove_na = TRUE) |>
     plot() -> p
  p + 
 labs(title = "Response to misinformation",
                   y ="Item"
                   )
```



How to summarize the variables in your dataset.

The easiest way is to use the table1 package.

```{r}
library(gtsummary)
mtcars |> tbl_summary()
```

Or if you just want to include some of the variables

```{r}
mtcars |> data_codebook(include = c(disp, hp, drat)) |>
     display()
```

```{r}
table1::table1(~., data = mtcars, 
               caption = "MT cars Variables") |> 
     t1flex() 

```

