---
title: "Cross_tabs2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cross_tabs2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(easystats)
library(ggplot2)
```

So far we've looked at how two variables are distributed
across every combination of values.

Howeve, we often are interested in how the two variables
relate to each other.  For example, are people who get
small coffees different than people who get large coffees?

This can be tricky because there are different numbers of 
orders in each group.

One way to understand this would be to break our table
into parts.


|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Small      | 10    |  9    | 19
|Medium     |  8    | 11    | 19
|Large      | 12    | 15    | 27

|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|
|Small      | 10    |  9    | 19

|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Medium     |  8    | 11    | 19


|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Large      | 12    | 15    | 27

Now we can calculate the percentages for each coffee size.

|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|
|Small      | 53%    |  47%    | 19

|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Medium     |  42%   | 58%    | 19

|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Large      | 44%    | 56%    | 27

It turns out the medium coffee group has the highest percentage
of orders with no milk.  This is true even though the large
group had more actual no milk orders.

Percentages are important because they let us make comparisons
on a _standardized_ basis.  That is we convert all of the 
rows to a representation that is based on having 100 observations
in the group. 

"Percent" comes from "Per cent" where cent means 100. 
Even though there are not 100 people in the group, we can still
calculate how many would be in each milk/no milk category if
the balance didn't change.

Because we calculated within the rows, we call these *row percents*.

We can put the table back together.


|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|
|Small      | 53%    |  47%    | 19
|-----------|-------|-------|----------
|Medium     |  42%   | 58%    | 19
|-----------|-------|-------|----------
|Large      | 44%    | 56%    | 27

We can tell that this table is row percents because adding 
up the values in each row give you 100.

Using the penguins data we can get row percents.

```{r eval=FALSE}
data_tabulate(penguins$year, by = penguins$species,
              proportions = "row",
              remove_na = TRUE) -> table
table |> print_md()
table |> plot() 
```


The percentage of penguins who are Adelie was pretty
stable over the three years. However, compared to 2007, in 2008 the 
percentage of penguins who are Chinstrap dropped 
substantially It then then recovered slightly in 2009. This
means that the percent who are Gentoo increased, since the 
percentagess have to add up to 100 within each row.

## Columns

We can do the same kind of process with columns.


|Size       |  Milk | No Milk  | Total
|-----------|-------|-------|----------
|Small      | 10    |  9    | 19
|Medium     |  8    | 11    | 19
|Large      | 12    | 15    | 27
|Total       | 30    |  35    | 65



|Size       |  Milk  
|-----------|-------
|Small      | 10     
|Medium     |  8     
|Large      | 12    
|           | 
|Total       | 30   


|Size       | No Milk   
|-----------|-------|
|Small      |  9    | 
|Medium     | 11    | 
|Large      | 15    | 
|           |       |
|Total       | 35    | 


Calculating the percentages base on the column totals


|Size       |  Milk  
|-----------|-------
|Small      | 33%     
|Medium     | 27%    
|Large      | 40%    
|           |   
|Total      | 30    


|Size       | No Milk   
|-----------|-------
|Small      | 26%     
|Medium     | 31%     
|Large      | 43%     
|           |
|Total       | 35     

These are the *column percents*.

Putting the table back together we can easily see how they
compare.

|Size       |  Milk |  No Milk
|-----------|-------|----------
|Small      | 33%    | 26%
|Medium     | 27%    | 31%
|Large      | 40%    | 43%
|           |      |   
Total       | 30   |  35

Looking at it from this perspective, we can see that 
people who ordered coffee without milk, generally 
ordered larger coffees than those who ordered with milk.

Now looking at the penguin data column percent, we can 
look at how year is distributed for each species.

```{r eval=FALSE}
data_tabulate(penguins$year, by = penguins$species,
              proportions = "column",
              remove_na = TRUE) -> table
table |> print_md()
table |> plot() 
```
