---
title: "Cross tabs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cross tabs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(easystats)
library(ggplot2)
```

```{r setup}
library(icorp26)
library(easystats)
```

"Cross tabs" are a kind of data presentation where two variables
are displayed together by "crossing" them. That means that 
the results for every combination of two variables are displayed.

For example, suppose there is a coffee shop which offers 
coffee with and without milk and in three sizes. 

There would be 6 ways of combining these two characteristics:

 - Small & Milk  
 - Small & No Milk  
 - Medium & Milk  
 - Medium & No Milk  
 - Large & Milk  
 - Large & No Milk  

You can rearrange the list however you want, but there always will
be 6 combinations, because $2*3=6$   The two comes from the
milk/no milk variable and the 3 from the three sizes.

Now suppose you have data on how many of each order type there
were on one day. 

|Combination         |  Number |
|--------------------|------   |
|  Small & Milk      | 10      |
|  Small & No Milk   |  9      |
|  Medium & Milk     |  8      |
|  Medium & No Milk  | 11      |
|  Large & Milk      | 12      |
|  Large & No Milk   | 15      |

Or you can rearrange the data to put the focus more on 
milk/no milk  rather than size.  

|Combination         |  Number |
|--------------------|------   |
|  Small & Milk      | 10      |
|  Medium & Milk     |  8      |
|  Large & Milk      | 12      |
|--------------------|---------|
|  Small & No Milk   |  9      |
|  Medium & No Milk  | 11      |
|  Large & No Milk   | 15      |


 If you add them up, you get $10+9 = 19$ that are small,
 $8+11 = 19$ that are medium, and $12+15 = 27$ that are large.
 
 Or you could say that you get $10+8+12 = 30$ that are with milk 
 and $9+11+15=25$ that are without milk.

You could also break this into two tables using one of
the variables. For example, we can make a separate set of tables
that break up the sample based on size. 


|Combination         |  Number |
|--------------------|------   |
|  Small & Milk      | 10      |
|  Small & No Milk   |  9      |

|Combination         |  Number |
|--------------------|------   |
|  Medium & Milk     |  8      |
|  Medium & No Milk  | 11      |

|Combination         |  Number |
|--------------------|------   |
|  Large & Milk      | 12      |
|  Large & No Milk   | 15      |

 
 You could have a question like: are some sizes of coffee
 more likely to be ordered with milk than others?
 
To answer this question you need to think about 
both variables at once.  A cross tab is one way to try
to answer ths kind of question.  Rearranging these by
"crossing" the variables, we get 3 rows and 2 columns (or 2
columns and 3 rows). By looking across the rows and down 
the columns we can understant each variable. And by comparing
rows to other rows and columns to other columns 

|Size       |  Milk | No Milk  |
|-----------|-------|-------|
|Small      | 10    |  9    |
|Medium     |  8    | 11    |
|Large      | 12    | 15    |

By reorganizing in this way we can more easily see how
each of the variables are distribute _and_ how they 
relate to each other. It is also a more compact layout.

Now let's look at some real data. We will use the 
penguins data set, which comes from a multiyear study
of penguins on three islands near the South Pole.

```{r}
data_tabulate(penguins$year, by = penguins$species, 
              remove_na = TRUE) -> table
   table |>  print_md()
```

Notice that in this cross tab the full sample distributions
are given in the `Total` column for year and the `Total` row 
for Island.

We can see that the biggest combination was 52 Adelie penguins
in 2009.
The smallest is 18 Chinstrap penguins in 2008.

We can show this graphically. In this kind of plot
the intensity of the color represents the size of the
numbers.

```{r}
table |> plot()
```


Another way to represent a cross tab is to convert the 
numbers to percentages.  

Looking at the coffee shop data we have 65 orders. 
If we take the number for a combination, divide by 65 and then
multiply the result by 100 we get the percents (rounded which 
means that they add up to 99 instead of 100).

|Combination         |  Number | Percentage
|--------------------|---------|-----------
|  Small & Milk      | 10      | 15%
|  Small & No Milk   |  9      | 14%
|  Medium & Milk     |  8      | 12%
|  Medium & No Milk  | 11      | 17%
|  Large & Milk      | 12      | 18%
|  Large & No Milk   | 15      | 23%

Again, this could be split into rows and columns. 

Graphically, we can display this view of the relationship
with a display where the size of the squares represents
the size of the percentages.

```{r}
data_tabulate(penguins$year, by = penguins$species,
              proportions = "full",
              remove_na = TRUE) 
```

```{r}
data_tabulate(penguins$year, by = penguins$species,
              proportions = "full",
              remove_na = TRUE) |> plot() |> print()
```


