---
title: "First analysis"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{First analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette illustrates some common things
you will want or need to do to get started.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

First we load the packages we need.
In R, packages are collections of useful data

```{r message=FALSE}
library(icorp26)
library(easystats)
library(here)
```

Let's say you want to use rss1.

[We can look at the codebook here.](https://www.cdc.gov/nchs/data/rss/codebook.pdf)

In R, one of the basic ways to refer to a variable is 
by combining the name of the dataset and the name of the variable
with a `$`. 

```
 dataset_name$variable_name
```


### If you get sick or have an accident, how worried are you that you will be able to pay yourmedical bills?

```{r}
data_tabulate(rss1$PAY_PAYWORRY)
```

### Overall Health
```{r}
data_tabulate(rss1$HIS_GENERAL)
```

### Marital Status
```{r}
data_tabulate(rss1$MARSTAT) 
```


Some things you should notice:

 - The N column tells you how many people answering the survey
fell into each category.

 - The `Raw %` Column includes everyone who was asked.

 - The `Valid %` Column includes only the people who answered
that specific question (not the <NA>s). 

 - At the top you can see: total N=7599 valid N=7557. The 
 difference of these two numbers (7599 - 7557) is 42 which
 is the same as the number of people have <NA>. 

- The `Cumulative %` adds each row to the total percent above it. See if you can follow the logic in the tables above.  This column is not relevant for variables that 
do not have some kind of ordering. Looking at our three
tables, the first two (how worried they are about losing
pay and overall self-reported health status _do_ have 
order) while the third, maritial status, does not. 

## A little bit more

R is very powerful in giving us many options for 
making our data easier to understand and our results
better looking.  Let's explore a bit.

### Marital Status

First, we can remove the missing data. In that
case the valid and complete data are the same.

```{r}
data_tabulate(rss1$MARSTAT, remove_na = TRUE) 
```

Then we can make our table look nicer by
using a customized "print" function.



```{r}
data_tabulate(rss1$MARSTAT, remove_na = TRUE) |> 
     print_md()

data_tabulate(rss1$MARSTAT, remove_na = TRUE) |> 
     print_html()

```

There are always a lot of different ways to do things
in R. The code chunk below is another way. 

When you search for help or try to use AI to help you
do things you will see many different approaches. 
That's fine; feel free to experiment. But always 
chek to make sure that the advice you get actually
works.


```{r}
report_table(rss1$MARSTAT, remove_na = TRUE) |> print_md()
```

```{r}
data_tabulate(rss1$P_INCOME_R, remove_na = TRUE) |>
     print_md()
```

We can see that income is given in a lot of categories
rather than exact number. But the categories are 
evenly spaced in $5000 steps until we get to the top
value of $150,000 or more. 

Because of this we cam treat this variable as though it
is a number and do things like calculate the mean and
median.

```{r}
mean(as.numeric(rss1$P_INCOME_R))
```

The value that cuts the sample into lower and upper halves 
is called the median.  The median here is 12. The 
12th category is "$75,000 to $84,999".  Look 
at the cumulative distribution in the first table
of income.   There are 7599 valid values If we divide
that by two we get 3799.5.  3754 values are in the 
categories  "$60,000 to $74,999" or lower. To
get to 3799 or 3800 we need to look in the next 
highest category.

```{r}
median(as.numeric(rss1$P_INCOME_R))
```
```{r}
summary(as.numeric(rss1$P_INCOME_R))
```

The summary shows the overall distribution of income,
including the highest and lowest values (Max and Min) and
the median and mean.  The 1st Quartile is the point 
the cuts off the bottom 25% while the 3rd Quartile
cuts off the bottom 75%. Look at the cumulative 
distribution and make sure you can see why. 

We can even do fancy things like combine results together
and get an automated paragraph (although there are not that
many options like this and you can look at its problems by
seeing what happens if you use `P_GENDERID` instead of
`P_GENDER`).

```{r}
participants <- 
 report_participants(
       rss1,
       age = NULL,
       sex = "P_GENDER",
       education = NULL,
       race = "P_RACE_R",
  ) 
```

`r participants`


There are many more things that you can do 

