---
title: "First analysis"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{First Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette illustrates some common things
you will want or need to do to get started.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

First we load the packages we need.

```{r}
library(icorp26)
library(easystats)
library(datawizard)
```

Let's say you want to use rss1.

[We can look at the codebook here.](https://www.cdc.gov/nchs/data/rss/codebook.pdf)

In R, one of the basic ways to refer to a variable is 
by combining the name of the dataset and the name of the variable
with a `$`. 

```
 dataset_name$variable_name
```


### If you get sick or have an accident, how worried are you that you will be able to pay yourmedical bills?

```{r}
data_tabulate(rss1$PAY_PAYWORRY)
```

### Overall Health
```{r}
data_tabulate(rss1$HIS_GENERAL)
```

### Marital Status
```{r}
data_tabulate(rss1$MARSTAT) 
```


Some things you should notice:

- The N column tells you how many people answering the survey
fell into each category.

- The `Raw %` Column includes everyone who was asked.

- The `Valid %` Column includes only the people who answered
that specific question (not the <NA>s). 

- The `Cumulative %` adds each row to the total percent above it. See if you can follow the logic in the tables above.





