---
title: "Introduction"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Your project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This summer we are using data from the Rapid Surveys
System that is part of the National Center for
Health Statistics. The NCHS is part of the 
Centers for Disease Control and Prevention.

[The RSS system](https://www.cdc.gov/nchs/rss/about.html) is designed to allow organizations
within CDC to quickly do topical surveys on emerging
issues. 

> Rapid Surveys can be used to:

> - Collect data on niche, emerging, or priority health topics
> - Close health information gaps
> - Learn how values and beliefs can affect access, care, and wellness
> - Assess questionnaire design to ensure data collection is meaningful
>   - Rapid Surveys are fielded several times each year using probability-based commercial online panels. Each survey includes a broad range of questions based on CDC’s needs at that time, so no two Rapid Surveys are alike.

The whole system is  well documented, with 
detailed codebooks and methodological information.

_Because of rapid change in government websites, make sure
that you download codebooks and other important
documentation._

The data are currently (May, 2026) provided 
as SAS data sets with input code.  To 
create the data sets in this package the data
sets were converted using the _haven_ package.
The simple converted data sets are included as 
`rss1_raw` through `rss7_raw`. 

For the data sets `rss1` through `rss7` more complex
data preparation steps were taken. 

- Missing data (all values with negative numbers) were converted to `NA` in R  

- SAS format information was extracted from the input files. Variables with formats were converted to factors.

- As of now, the ordered attribute for the factors is not
included. However labels should be in the correct order.

## Technical caution

As explained in the documentation, correct estimates
of standard errors requires that sample weights be 
used. However, the weights are not part of the public
data sets.  As a result, estimates of standard
errors are smaller than what they should be, leading
to increased risk of incorrect results of traditional hypothesis
tests. That is, results will appear to be
statistically significant when there is no 
difference in the population. The null hypothesis
will be rejected incorrectly.

For the SROP we will mainly deal with this by using 
a more stringent criteria to decide whether the 
results are "statistically significant." If you have taken
statistics before, you can think of this as saying
we change the p value for determining if we can reject a
null hypothesis from .05 to .025.  If you don't know
what this means, don't worry about it. We will discuss
this more later in the summer. 

Another thing we will do is pay more attention to 
effect size than to the p values. 

If you have taken statistics before you may want to
try some more complex or sophisticated
approaches that are not covered in the workshop that is
great! Feel free to experiment.

If you really want to go deep, 
it may be that Multilevel Regression Post Stratification (Mr P)
could be used to address this issue.  See:

- [Mr. P Case Studies](https://bookdown.org/jl5522/MRP-case-studies/introduction-to-mrp.html)  
 
- [Multilevel Regression with Post Stratifcation](https://en.wikipedia.org/wiki/Multilevel_regression_with_poststratification) 

However, for the summer program we are mainly  going to 
ignore this issue and focus on other methods,

## References

CDC. 2026. “NCHS Rapid Surveys System.” https://www.cdc.gov/nchs/rapid-surveys/about/index.html.

Wickham H, Miller E, Smith D (2025). _haven:
  Import and Export 'SPSS', 'Stata' and 'SAS'
  Files_. doi:10.32614/CRAN.package.haven
  <https://doi.org/10.32614/CRAN.package.haven>,
  R package version 2.5.5,
  <https://CRAN.R-project.org/package=haven>.
  

```{r eval=FALSE}
citation("haven")
```

