---
title: "Significance tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Significance tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Work in Progress

```{r setup}
library(icorp26)
library(easystats)
```

So far we have been looking at statistics that describe 
people in the RSS sample. 
However, in the social sciences we are almost always interested
in whether we can draw any conclusions (make an _inference about_)
the larger population.

The _sample_ is drawn from a larger population. In the case of 
the RSS the populaton of interest is the whole adult population
of the United States, about 253 million people. The rss1 data
set has 7599 observations. That's a big number, but just 0.003%
of  the total population. 

Even with the very best samplng procedures it is extremely unlikely that 
a sample will match the population exactly. And it is possible
that a sample could be very inaccurate.

Statisticians have come up with many different ways to 
address this problem. Some of these involve _how the sample is
selected_. Others involve how to calculate statistics that 
are based on samples instead of populations. A third
approach focuses on the math of how samples relate to populations.
It turns out that there are patterns in this and we can leverage
them. Over the last 200 years statisticians have developed
several approaches to this.

One approach to  this is called Null Hypothesis 



