---
title: "Significance tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{significancetests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```



Work in Progress

```{r setup}
library(icorp26)
library(easystats)
```

So far we have been looking at statistics that describe 
people in the RSS sample. 
However, in the social sciences we are almost always interested
in whether we can draw any conclusions (make an _inference about_)
the larger population.

The _sample_ is drawn from a population. In the case of 
the RSS the populaton of interest is the whole adult population
of the United States, about 253 million people. The rss1 data
set has 7599 observations. That's a big number, but just 0.003%
of  the total population. 

Why would we think we can use such a small number of people to
represent a whole population?

Even with the very best samplng procedures it is extremely unlikely that 
a sample will match the population _exactly_. And it is possible
that a sample could be very inaccurate.

Let's look at why that is using Statkey

https://www.lock5stat.com/StatKey/

What is interesting and useful is that over _many_ samples there are 
certain patterns that emerge in the results. 



Statisticians have come up with many different ways to 
address this problem. Some of these involve _how the sample is
selected_. Others involve how to calculate statistics that 
are based on samples instead of populations. A third
approach focuses on the math of how samples relate to populations.
It turns out that there are patterns in this and we can leverage
them. Over the last 200 years statisticians have developed
several approaches to this.

One approach to  this is called Null Hypothesis Significance Testing. This is usually the first approach
that students learn, and it is the one you most often 
see in publications, especially older ones.

Some of the most commonly used NHST are the t test, the F test and 
the Chi Square test. The choice of tests depends on the variables
and assumptions we are willing to make. However they all follow the 
same logic.

A null hypothesis takes the form:

There is no relationship between variable A and variable B.

Here are some examples of what this could mean:

The mean of sun screen misinformation is the same
for all racial and ethnic groups.  

The proportion of people in our population of interest
who use sunscreen most of the time is
the same for all racial/ethnic groups.  

The "alternative hypothesis" is that there _is_ a
relationship between the variables in the population.

*Challenge*: When we take random samples of a population
the relationships in the _sample_ may not match what is in
the _population._  



## Step one 



## Important points to remember

1. Statistically significant does not mean *important*.  
2. We never accept hypotheses, we just either reject or fail to 
reject the _null hypothesis_ of no difference.  
3. Tests assume random sampling with replacement. 
4. Statistical significance is _not_ about the probability that a hypothesis is true.







