---
title: "two_variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{two_variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(icorp26)
library(easystats)
library(here)
```

Most of the time in research we are interested 
in understanding the relationships between variables.
That is, we want to know whether differences in one
variable are associated with differences in another
variable.

Let's look at whether there is a relationship between
self-reported health and how worried respondents are
about losing pay if they become sick.

We can see that both variables are ordinal, meaning 
that there are distinct categories that have a logical
ordering.

We can also think of the health status as the independent
variable and worry as the dependent variable. 
For example, we might think that those with worse 
health would be more worried


```{r}
data_tabulate(rss1$PAY_PAYWORRY, by = rss1$HIS_GENERAL)
```

We can make this more readable by dropping the 
missing cases and using a nice print function.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE,
              name = "Worry by health status") |> 
     print_html()
```

To make it easier to interpret we can add the 
column percentages to the table. We always 
calculate the percents within categories of 
the independent variable.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE,
              proportions = "column",
              name = "Worry aboout pay") |> 
     print_html()
```

Looking at the percents, we can see that those who 
are in excellent health are generally less worried than
those in poor health.  This doesn't apply to everyone--
there are people in poor health who are not worried and 
people in good health who are worried--but it is 
an overall pattern.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE)  |> 
     effectsize()
```

```{r}
# group the data by your control variable
rss1 |> data_group("P_GENDER") -> pg
data_tabulate(pg, "PAY_PAYWORRY", 
              by = "HIS_GENERAL",
              remove_na = TRUE) -> tableobject
tableobject 
tableobject |>
     data_chisq() 
     
 
```

```{r}
rss1 |> 
     data_tabulate(x = rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE)
```

