---
title: "Analysis of two_variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis of two_variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(icorp26)
library(easystats)
library(here)
```

Most of the time in research we are interested 
in understanding the relationships between variables.
That is, we want to know whether differences in one
variable are associated with differences in another
variable.

Let's look at whether there is a relationship between
self-reported health and how worried respondents are
about losing pay if they become sick.

We can see that both variables are ordinal, meaning 
that there are distinct categories that have a logical
ordering.

We can also think of the health status as the independent
variable and worry as the dependent variable. 
For example, we might think that those with worse 
health would be more worried


```{r}
data_tabulate(rss1$PAY_PAYWORRY, by = rss1$HIS_GENERAL)
```

We can make this more readable by dropping the 
missing cases and using a nice print function.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE,
              name = "Worry by health status") |> 
     print_html()
```

To make it easier to interpret we can add the 
column percentages to the table. We always 
calculate the percents within categories of 
the independent variable.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE,
              proportions = "column",
              name = "Worry aboout pay") |> 
     print_html()
```

Looking at the percents, we can see that those who 
are in excellent health are generally less worried than
those in poor health (7.5% compared
to 40.6%). And the middle income groups are between these  This doesn't apply to everyone--
there are people in poor health who are not worried and 
people in good health who are worried--but it is 
an _overall pattern_.

One way that researchers use to talk 
about how strong the relationship is uses
the concept of "effect size." 
An effect size of 0 would mean that 
the two varables are totally unrelated 
(the column percents would be the same
for all of the columns in our table).
An effect size of 1 would mean that we 
know 100% of the time what value
the people in a column have, because 
they all have the same value. 

Cramer's V is one way to measure 
effect size.  For our table, the 
effect size is .19. This is only 
based on our sample, so the confidence
interval gives us a general idea of 
what we could say about the population
the sample is drawn from. In this
case the potential values are pretty 
broad, somewhere between .18 and 1.0.

```{r}
data_tabulate(rss1$PAY_PAYWORRY, 
              by = rss1$HIS_GENERAL,
              remove_na = TRUE)  |> 
     effectsize()
```

Another thing that researchers usually do
is to test for "statistical significance."

Statistical significance asks "if the true
effect size _in the population_ is 0, how
likely is it that a random sample of 
the population would have a relationship 
the size of the relationship in our sample 
or even stronger. 

That is a lot! And the concept is complicated.
Still, as a rule of thumb, many analysts say
that if the probability is less than 5% (or p < .05)
that they are comfortable saying they "reject
the null hypothesis" that there is no 
relationship between the variables.

A common way of doing this for two 
categorical (nominal) variables is with
the chi square test.

```{r}
# group the data by your control variable

data_tabulate(rss1, "PAY_PAYWORRY", 
              by = "HIS_GENERAL",
              remove_na = TRUE) |> 
     data_chisq() 
     
 
```

Because the p-value is less than .05 
(< 2.2e-16) convention says we can reject the 
null hypothesis.  

In case it has been a while since you used 
scientific notation the `e-16` means we could
translate that number to:

.000000000000000022  

which is definitely less that .05!



In this case because the variables 
are both ordered we could also use a test
(and effect size measure) such as Kendall's tau
if we convert them to numbers.  This requires
some assumptions, but we can see 

```{r}
cor.test(as.numeric(rss1$PAY_PAYWORRY),  as.numeric(rss1$HIS_GENERAL),
                            method = "kendall", 
                            alternative = "two.sided")
```


In this analysis the value of tau is -0.219.
How can it be negative? Weren't they supposed
to be between 0 and 1?  The answer is
that the absolute size is between 0 and 1
if you take the absolute value, which means
turning a negative number into a positive number.  
So the effect size using tau is slightly 
larger than that using Cramer's V.

The negative sign gives us some additional
information about the _direction_ of the
relationship.  In this case, as the value
of one variable goes up, the value of the 
other variable goes down. So the 
higher the score on health status, the lower
the score on worry. What is going on? 

If we look at a table of how the numeric 
and non numeric varsions of GENERAL_HIS
related to each other, we see that people
in *poor* health were scored with a 5, 
while people in *excellent* health 
were scored with a 1.  So the higher
someone's score, the worse their health.

When we do the same for PAY_PAYWORRY, 
the people who were very worried got 
a 1 and the not at all worried got a 3.
So the higher the scor, the less the 
worry.

So, as we would expect, the negative sign
indicates that as self-reported health 
gets worse, worry about pay gets worse.

This is why it is very important to always 
make sure you understand your variables before
working with them or interpreting the results.

