---
title: "Knowledge of sunburn risks"
author:
  - name: "Elin Waring"
    affil: 1
affiliation:
  - num: 1
    address: Dept. of Sociology, Lehman College
column_numbers: 3 
logoright_name: https&#58;//raw.githubusercontent.com/elinw/icorp26/refs/heads/main/inst/images/LehmanLogo_White_Vertical.png
logoleft_name: https&#58;//raw.githubusercontent.com/elinw/icorp26/c34a9b74b3fbd736b667872fdc52af1767f56654/inst/images/icorplogo.jpg
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: packages.bib
primary_colour: "#00365B"
secondary_colour: "#ffffff"
accent_colour: "#0061A0"
sectitle_bgcol: "#00365B"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```
```{r}
library(easystats)
library(icorp26)
library(ggplot2)
```

# Introduction

Skin cancer is the most common cancer in the United States.
According to the Centers for Disease Control "Most skin cancers are caused by too much exposure to ultraviolet (UV) light. Protection from UV rays is important all year." (CDC) More than 
1 in every 5 Americans experience at least one diagnosis of skin cancer by the age of 70. 
There are major differences in the rates and types of skin cancer
and location of the cancer across racial and ethnic groups. Hispanic and Black Americans 
are more likely to be diagnosed at late stages of skin cancer than
white Americans (Skin Cancer Foundation).

There are known effective ways to prevent sun cancer, but these
often require individuals to take various steps to avoid exposure 
of their skin to to UV rays. These include covering the skin,
correct use of effective sunscreens and not using tanning beds.

Although people of all races get skin cancer, reseearchers have found
that medical text books and educational materials for the public often focus exclusively on "fair" (i.e. white) skin (Rodini and Kowalsky 2021). This may lead both doctors and patients to 
ignore the risk of UV rays to to those with darker skin.


## Objectives

1. To determine the relationship of 


# Methods

The RSS surveys are topical surveys administered to US adults. 
The first round of RSS included a set of questions about sun bur risks
and sun screen use. 

Hypothesis 1: People who have more _incorrect_ information
about sun burn will use sunscreen less often than those with
correct information.  

Rationale: Using sunscreen requires taking an action and 
also expenses. If someone does not think sun burn is a problem, 
they won't make the effort or spend the money.

Hypothesis 2: People racialized as white will be less likely
to agree with misinformation than those racialized as Black or Hispanic.

Rationale: Racialized information on skin cancer, emphasizing 
the risks of fair skin, leads both the general population
and doctors to minimize the risks among people with 
darker skin. Therefore we expect that groups that on average have
darker skin will tend to have less accurate information on the 
risks of sunburn.

# Results

The RSS data on know

```{r eval=FALSE}
data_tabulate(c(rss1$SUN_CLOUDY), by =  rss1$P_RACE_R, proportions = "column", remove_na = TRUE)
data_tabulate(c(rss1$SUN_NOHARM), by =  rss1$P_RACE_R, proportions = "column", remove_na = TRUE)

data_tabulate(rss1$SUN_USESKIN, by =  rss1$SUN_CLOUDY, proportions = "column", remove_na = TRUE)
data_tabulate(rss1$SUN_USESKIN, by =  rss1$SUN_NOHARM, proportions = "column", remove_na = TRUE)
```
```{r}
# Students: Notice that here I realized that to have the graph make sense
# I had to say that the group variable was numeric because it 
# didn't automatically go in the right order in the graph.
# Also to get a mean I needed to make the dependent variable
# numeric

means_by_group(to_numeric(rss1$SUN_USESKIN), by= to_numeric(rss1$SUN_NOHARM)) |>
     plot() + labs(title = "Use of Sunscreen by Misinformation (SUN_NOHARM", xlab = "Ageement that there is no long term risk")
```
```{r}
means_by_group(to_numeric(rss1$SUN_USESKIN), by=to_numeric(rss1$SUN_CLOUDY)) |>
     plot() + labs(title = "Use of Sunscreen by Misinformation (SUN_CLOUDY", xlab = "Ageement that there is no risk")

```

Overall Hypothesis 1 is generally supported. For example,
those who agreed more that there is no risk from the sun 
on cloudy days uses sun screen less often and the result was statistically
significant at the .001 level. 

However the results for disagreeing that "sunburn is not harmful in long run"
are less clear. Those who "strongly agree" with the misinformation 
have similar sun screen use to those who "strongly disagree." 
The unexpectedly high sun screen use for this group should be invetigated further.

```{r}
means_by_group( to_numeric(rss1$SUN_NOHARM), by =rss1$P_RACE_R) |>
     plot() + labs(title = "Sunscreen Misinformation by Race/Ethnicity (SUN_NOHARM", xlab = "Ageement that there is no long term risk")

```
```{r}

means_by_group(to_numeric(rss1$SUN_CLOUDY), by =rss1$P_RACE_R) |>
     plot() + labs(title = "Sunscreen Misinformation by Race (SUN_CLOUDY", xlab = "Ageement that there is no risk")
```


# Next Steps

The next steps for this research will be to incorporate 
other variables, such as age, level of education, language and
whether a person has a doctor. Research using other data sets has found that people get most of their health information
from lay sources (Ford and Kaphingst 2009), including 
community organizations, friends, family and the media. They
found that this related to beliefs that the risks of cancer 
are modifiable. There are some items in the RSS that might
allow examination of this. 

The issue of the unexpected results for the "strongly agree" 
group requires further investigation.

# Conclusion

Overall, these findings support sociological research that has 
identified ways in which medical information is almost always
shown in reference to white bodies. While the need for education
about the dangers of UV rays is clear, success at reducing skin
cancer among African Americans, and to a lesser extent Hispanics, will require
addressing the structural issues of a racialized health care system.


# Key References

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```


